---
title: "Project"
author: "Anthony Yun avyun@umich.edu"
date: "2023-04-08"
output: pdf_document
---

## SETUP
I still need to make error values

```{r}
library(haven)
library(tidyverse)
library(glmnet)
body <- read_xpt('P_BMX.XPT')
exam <- read_xpt('P_DEMO.XPT')

dietary <- read_xpt('P_DR1IFF.XPT')
oralHealth <- read_xpt('P_OHXREF.XPT')
```

```{r}
oralHealthSub = subset(oralHealth, select = c("SEQN", "OHAROCDT"))
oralAndFood = merge(oralHealthSub, dietary, by = "SEQN")
oralAndFood[["SEQN"]] = NULL
oralAndFood = na.omit(oralAndFood)

#Remove "Refused" and "Don't Know" data points
oralAndFood = filter(oralAndFood, DR1_040Z != 9)
oralAndFood = filter(oralAndFood, DR1_040Z != 7)
oralAndFood = filter(oralAndFood, DR1FS != 99)
oralAndFood = filter(oralAndFood, DR1_030Z != 99)
oralAndFood = filter(oralAndFood, DR1DRSTZ != 5)
```


```{r}
## SETUP FOR RIDGE AND LASSO
set.seed(1)
X = model.matrix(OHAROCDT ~ ., data = oralAndFood)[, -1]
decay = oralAndFood$OHAROCDT
grid <- 10^seq(10, -2, length = 100)
ridge.mod <- glmnet(X, decay, alpha = 0, lambda = grid)
train_idx <- sample(1:nrow(X), floor(.7*nrow(X)), replace = FALSE)

trainingSet <- oralAndFood[train_idx,]
testingSet <- oralAndFood[-train_idx,]
```

\break 

## RIDGE

```{r}
## RIDGE 
set.seed(1)
cv.out <- cv.glmnet(X[train_idx,], decay[train_idx], alpha = 0)
bestlam <- cv.out$lambda.min
bestlam
ridge.mod <- glmnet(X, decay, alpha = 0, lambda = grid)
ridge.pred <- predict(ridge.mod, s = bestlam, newx = X[-train_idx,])
out <- glmnet(X, decay, alpha = 0)
predict(out, type = "coefficients", s = bestlam)[1:84,]
```

```{r}
## Still need to figure out how to make the True/False table for the ridge regression instead
ridge.pred.train = predict(ridge.mod, newx = X[train_idx, ])
ridge_train_error = mean((ridge.pred.train - decay[train_idx])^2)
ridge_train_error

ridge.pred.test = predict(ridge.mod, newx = X[-train_idx, ])
ridge_test_error = mean((ridge.pred.test - decay[-train_idx])^2)
ridge_test_error
```

## LASSO

```{r}
## LASSO
set.seed(1)
cv.out <- cv.glmnet(X[train_idx,], decay[train_idx], alpha = 0)
bestlam <- cv.out$lambda.min
bestlam
lasso.mod <- glmnet(X[train_idx,], decay[train_idx], alpha = 1)
lasso.pred <- predict(lasso.mod, s = bestlam, newx = X[-train_idx,])
out <- glmnet(X, decay, alpha = 1)
predict(out, type = "coefficients", s = .001)[1:84,]
```

Important variables (not zero coeficent): DR1DBIH, DR1DAY, DR1CCMNM, DR1FS, DR1IPFAT, DR1ISODI


```{r}
#lasso.pred.train = predict(lasso.mod, newx = X[train_idx, ])
#lasso_train_error = mean((lasso.pred.train - decay[train_idx])^2)
#lasso_train_error
#lasso.pred.test = predict(lasso.mod, newx = X[-train_idx, ])
#lasso_test_error = mean((lasso.pred.test - decay[-train_idx])^2)
#lasso_test_error
```

```{r}
## Using chosen lasso model predictors

modLasso <- lm(OHAROCDT~DR1DBIH+ DR1DAY+ DR1CCMNM+ DR1FS+ DR1IPFAT+ DR1ISODI, data = trainingSet)

train_pred_lasso <- predict(modLasso, newdata = trainingSet)
test_pred_lasso <- predict(modLasso, newdata = testingSet)

train_tab_lasso <- table(trainingSet$OHAROCDT, train_pred_lasso > 0.5)
test_tab_lasso <- table(testingSet$OHAROCDT, test_pred_lasso > 0.5)

train_tab_lasso
test_tab_lasso
table(trainingSet$OHAROCDT)
table(testingSet$OHAROCDT)
```
Perfect accuracy 

\break

## Forward STEP 

```{r}
# Forward step
library(SignifReg)
library(leaps)

nullmodel <- lm(OHAROCDT ~ 1, data = oralAndFood)
fullmodel <- lm(OHAROCDT ~ ., data = oralAndFood)

## This will take a a few minutes
select.p.fwd <- SignifReg(fit = nullmodel, scope = list(lower = formula(nullmodel), upper = formula(fullmodel)),alpha = 0.05, direction = "forward", adjust.method = "none", trace = FALSE)
summary(select.p.fwd)
```

Important variables: DR1DAY, DR1FS, DR1FS, DR1DBIH, DR1ISODI, DR1CCMNM, DR1ITHEO, DR1IFDCD, DR1IFIBE DR1IPFAT, DR1ICALC, DR1IS180

Common variables with Lasso:  DR1DBIH, DR1DAY, DR1CCMNM, DR1FS, DR1IPFAT, DR1ISODI (all of the lasso ones)


```{r}
#using that model

modFwd <- lm(OHAROCDT ~ DR1DAY + DR1FS + DR1FS + DR1DBIH + DR1ISODI + DR1CCMNM + DR1ITHEO + DR1IFDCD + DR1IFIBE + DR1IPFAT+ DR1ICALC + DR1IS180, data = trainingSet)

train_pred_fwd <- predict(modFwd, newdata = trainingSet)
test_pred_fwd <- predict(modFwd, newdata = testingSet)

train_tab <- table(trainingSet$OHAROCDT, train_pred_fwd > 0.9)
test_tab <- table(testingSet$OHAROCDT, test_pred_fwd > 0.9)

train_tab
test_tab
table(trainingSet$OHAROCDT)
table(testingSet$OHAROCDT)
```

We see perfect training and testing accuracy from confusion table.

```{r}
#train_error <- mean((trainingSet$OHAROCDT - train_pred_fwd)^2)
#test_error <- mean((testingSet$OHAROCDT - test_pred_fwd)^2)
#train_error
#test_error
```